Non-word repetition in Yélî Dnye

### 2020-10-05 manuscript.Rmd
version that knits for sure

@@

<!-- M2A: regarding demo, it looks like you have more mateds than birth orders because family data was systematically not collected for Pu. kids and was missing for three kids tested in No. who were actually from other hamlets. New xslx files sent via email! I left age recomputing to you since I was confused about which file would be used.-->
<!-- a2m thanks!!! now inc'ed -->

<!-- M2A: regarding NAs in stim_seg_freq, these appear for 'lv' stimuli because, in the 'segments' table, freq_corpus and freq_corpus_types are NA for that grapheme). The 'segments' table has NAs for 'lv' because it is not included in segment-counts.txt or segment-counts-types.txt. Note that that these NAs are excluded when adding mean item frequency to the 'stims' table (wrangling.R/l.136 & l.145), but this has a pretty big impact since those are all two-segment words. Not sure how avoidable this is, going back to our discussion before about not wanting to ad arbitrarily small frequencies. Thoughts? -->
<!-- a2m So your fear is that frequency is overestimated in these cases, right? Perhaps we can assign the missing segments the lowest observed frequencies? -->

@@

### The Yélî community
<!--M2A: is this okay?-->
<!-- a2m I think it's great! although it seemed a bit unexpected to explain all of this, so I added some justification -->

@@

read.csv("nwr-demo.csv",header=T,sep=",")->demo
#colnames(demo)

#lines no longer needed
#if family is missing, attribute each child to a different one
#demo$familyID=as.character(demo$familyID)
#demo$familyID[is.na(demo$familyID)]<-paste("k",demo$ID[is.na(demo$familyID)])
#recalc age
#demo$age.rounded=as.numeric(difftime(as.Date(as.character(demo$Date),format="%Y%m%d"),as.Date(as.character(demo$DOB.ISO),format="%Y%m%d"),unit="weeks"))/52.25
write.table(demo,"nwr-demo.csv")

@@

We could include `r summary(xl_fr)$devcomp$dims["N"]` observations, from `r summary(xl_fr)$ngrps["file"]` children producing in any given trial one of `r summary(xl_fr)$ngrps["target"]` potential target words. The analysis revealed a main effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p = `r mycoeff["age.rounded_s","Pr(>|z|)"]`); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr_s","Estimate"]`, SE ß = `r mycoeff["avg_fr_s","Std. Error"]`, p = `r mycoeff["avg_fr_s","Pr(>|z|)"]`): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r r(`r length_r$parameter`) = `r round(xlfr_r$estimate,3)`.

@@

[^2]: We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less convincing.

@@

xl_fr=glmer(correct~avg_fr_s+age.rounded_s+(1|file) + (1|target),data=trnsc,family="binomial")
mycoeff=round(summary(xl_fr)$coefficients,3)

@@

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`.
<!-- a2m it doesn't make sense to do this, right, with only 3 items? -->
<!-- M2A: you mean with only three lengths? I guess I would prefer a little regression -->

>>>>

## TODO Alex

- check again if all fields are explained
- set up OSF (or revise if already there)
- implement decisions on prop phones & patterns nwr mp
- revise bullet points for discussion

@@

@@

### The Yélî community

@@

read.csv("nwr-demo.csv",header=T)->demo
#colnames(demo)

#lines no longer needed
#if family is missing, attribute each child to a different one
#demo$familyID=as.character(demo$familyID)
#demo$familyID[is.na(demo$familyID)]<-paste("k",demo$ID[is.na(demo$familyID)])
#recalc age
#demo$age.rounded=as.numeric(difftime(as.Date(as.character(demo$Date),format="%Y%m%d"),as.Date(as.character(demo$DOB.ISO),format="%Y%m%d"),unit="weeks"))/52.25
#write.table(demo,"nwr-demo.csv",row.names=F,quote=T,sep=",")

@@

We could include `r summary(xl_fr)$devcomp$dims["N"]` observations, from `r summary(xl_fr)$ngrps["file"]` children producing in any given trial one of `r summary(xl_fr)$ngrps["target"]` potential target words. The analysis revealed a main effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p = `r mycoeff["age.rounded_s","Pr(>|z|)"]`); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr_s","Estimate"]`, SE ß = `r mycoeff["avg_fr_s","Std. Error"]`, p = `r mycoeff["avg_fr_s","Pr(>|z|)"]`): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r r(`r xlfr_r$parameter`) = `r round(xlfr_r$estimate,3)`.

@@

[^2]: We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.

@@

len_lm=glmer(correct~type+age.rounded_s+(1|file) + (1|target),data=trnsc,family="binomial",subset=c(trnsc$type!=1))
mycoeff=round(summary(len_lm)$coefficients,3)

@@

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`. In a generalized binomial mixed model, we included `r summary(len_lm)$devcomp$dims["N"]` observations, from `r summary(len_lm)$ngrps["file"]` children producing in any given trial one of `r summary(len_lm)$ngrps["target"]` potential target words. The analysis revealed a main effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p = `r mycoeff["age.rounded_s","Pr(>|z|)"]`); and a significant estimate for the length of the target words in number of syllables (ß = `r mycoeff["type","Estimate"]`, SE ß = `r mycoeff["type","Std. Error"]`, p = `r mycoeff["type","Pr(>|z|)"]`).
